---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filestamp <img src="man/figures/filestamp-animated-logo.svg" width="170" align="right" alt="A hexagonal logo for filestamp"/> 

<!-- badges: start -->
[![R-CMD-check](https://github.com/coatless-rpkg/filestamp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coatless-rpkg/filestamp/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`filestamp` makes it easy to add and maintain consistent headers across all files in your project, regardless of programming language. Headers can include copyright notices, file descriptions, authorship information, and other metadata.

## Installation

You can install the development version of `filestamp` from [GitHub][filestamp-gh] with:

```{r}
#| label: setup
#| eval: false
# install.packages("remotes")

# From GitHub
remotes::install_github("coatless-rpkg/filestamp")
```

## Getting Started

This section provides a few examples of how to use `filestamp` to add headers to your files. For more detailed information, please refer to the package documentation.

### Loading the Package

To use `filestamp`, you first need to load the package using the `library()` function. This will make all the functions and features of `filestamp` available for use in your R session.

```{r}
#| label: setup-usage
library(filestamp)
```

### Stamping a Single File

You can use the `stamp_file()` function to add a header to a single file. The function takes the file path as an argument and can also accept a template name or a custom template.

```{r}
#| label: add-headers
#| eval: false

# Stamp a file with the default template
stamp_file("script.R")

# Stamp a file with a specific template
stamp_file("script.py", template = "mit")

# Preview changes without modifying the file
stamp_file("script.R", action = "dryrun")

# Create a backup before stamping
stamp_file("important_script.R", action = "backup")
```


### Stamping a Directory with Multiple Files

You can use the `stamp_dir()` function to add headers to all files in a directory. You can specify a pattern to match specific file types, and you can also choose to stamp files in subdirectories.

```{r}
#| label: stamp-dir
#| eval: false
# Stamp all R files in a directory
stamp_dir("src", pattern = "\\.R$")

# Stamp all files in a directory and subdirectories
stamp_dir("project", recursive = TRUE)

# Stamp with a specific template
stamp_dir("src", template = "gpl-3")
```

## Templates

### Using Built-in Templates

The package comes with several built-in templates:

- `default` - A simple header with copyright, author, and license information
- `mit` - MIT License header
- `gpl-3` - GNU General Public License header

List available templates:


```{r}
#| label: stamp-templates
stamp_templates()
```

### Creating Custom Templates

You can create your own custom templates using the `stamp_template_create()` function. This allows you to define the fields and content of the header.

```{r}
#| label: custom-template
#| eval: false
# Create a custom template
my_template <- stamp_template_create(
  name = "my_custom",
  fields = stamp_template_describe(
    copyright = stamp_template_field("copyright", "MyCompany 2025", required = TRUE),
    author = stamp_template_field("author", "John Doe", required = TRUE)
  ),
  content = stamp_template_content(
    "Copyright (c) {{copyright}}",
    "Created by: {{author}}",
    "File: {{filename}}"
  )
)

# Use the custom template
stamp_file("script.R", template = my_template)
```

## Updating Existing Headers

You can update existing headers in files using the `stamp_update()` function. This is useful for changing copyright years, adding new authors, or modifying other metadata.

> [!IMPORTANT]
>
> Make sure to use the `stamp_update()` function carefully, as it will modify existing files. Always create backups before making changes.
>
> Some helper functions may not work as expected if the file does not contain a header
> or if the header is not in the expected format.

```{r}
#| label: existing-header
#| eval: false
# Update copyright years in existing headers
stamp_update("old_script.R", list(
  copyright = stamp_update_helper_copyright_extend()
))

# Add a new author to existing headers
stamp_update("collaborative_script.R", list(
  author = stamp_update_helper_author_add("Jane Smith")
))
```

## Language Support

The package supports many programming languages out of the box:

```{r}
#| label: language-supported
#| eval: false
# List supported languages
languages()

# Register a custom language
language_register(
  "kotlin",
  extensions = c("kt", "kts"),
  comment_single = "//",
  comment_multi_start = "/*",
  comment_multi_end = "*/"
)
```

## Variables

You can use variables in your templates to customize the header content. The package provides several built-in variables, and you can also define your own.

```{r}
#| label: stamp-template-variables
#| eval: false
# List available variables
stamp_variables_list()

# Add a custom variable
stamp_variables_add("team", "Data Science")

# Set company name
options(filestamp.company = "Acme Corp")
```

## License

AGPL (>= 3)

[filestamp-gh]: https://github.com/coatless-rpkg/filestamp
