---
title: "Maintaining headers"
vignette: >
  %\VignetteIndexEntry{Maintaining headers}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{=html}
<style>
.fs-fig{display:block;margin:1.5rem auto;max-width:100%;height:auto}
.fs-dark{display:none}
@media (prefers-color-scheme:dark){.fs-light{display:none}.fs-dark{display:block}}
[data-bs-theme=light] .fs-light{display:block}
[data-bs-theme=light] .fs-dark{display:none}
[data-bs-theme=dark] .fs-light{display:none}
[data-bs-theme=dark] .fs-dark{display:block}
</style>
```

```{r setup}
#| include: false
library(filestamp)
```

Headers age. Copyright years roll over and contributors join. `stamp_update()`
edits an existing header in place, changing only the fields you name and leaving
the rest of the file alone.

![stamp_update revises an existing header in place, bumping the copyright year and adding an author while leaving everything else unchanged.](../man/figures/update-flow-light.svg){.fs-fig .fs-light}
![](../man/figures/update-flow-dark.svg){.fs-fig .fs-dark aria-hidden="true"}

## The common cases

Two verbs cover the everyday tasks. Here is a file with a header from a few
years ago:

```{r}
f <- tempfile(fileext = ".R")
writeLines(c(
  "# Copyright (c) 2020",
  "# Author: Jane Doe",
  "",
  "x <- 1"
), f)
```

`stamp_bump_year()` extends the copyright to the current year, keeping the owner
and the start year. `stamp_add_author()` adds a contributor without duplicating
anyone already listed:

```{r}
#| message: false
stamp_bump_year(f)
stamp_add_author(f, "Sam Smith")
cat(readLines(f)[1:2], sep = "\n")
```

Running either again is safe: the year is already current and Sam is already
listed, so nothing changes.

## Custom updates

For anything beyond those two, `stamp_update()` takes named updates directly.
Each value is a new value, or a function of the field's current value.
`year_extend()` and `author_add()` build those functions, and you can mix them
with plain values in a single call:

```{r}
#| message: false
g <- tempfile(fileext = ".R")
writeLines(c(
  "# Copyright (c) 2018",
  "# Author: Ada Lovelace",
  "# License: GPL-2",
  "",
  "y <- 2"
), g)

stamp_update(g,
  copyright = year_extend(),
  author    = author_add("Sam Smith"),
  license   = "MIT"
)
cat(readLines(g)[1:3], sep = "\n")
```

## Reuse a set of edits across files

`stamp_edits()` and `stamp_update()` do different jobs. `stamp_edits()` builds a
reusable *bundle* of changes and touches no files on its own. `stamp_update()`
is what *applies* changes to files, whether you pass them inline or as a bundle.

Bundling pays off when you want the same set of edits applied in more than one
place. The bundle also prints what it will change:

```{r}
edits <- stamp_edits(
  copyright = year_extend(),
  author    = author_add("Sam Smith")
)
edits
```

## Update a whole directory

`stamp_update()` and the verbs accept a directory as well as a file. Given a
directory, they update every file that already has a header and leave the rest
alone. Use `recursive = TRUE` to include subdirectories.

![A reusable edit bundle is applied to every already-stamped file in the tree, extending each file's copyright year; a file with no header is skipped.](../man/figures/update-directory-light.svg){.fs-fig .fs-light}
![](../man/figures/update-directory-dark.svg){.fs-fig .fs-dark aria-hidden="true"}

```{r}
#| message: false
project <- file.path(tempdir(), "project")
dir.create(file.path(project, "R"), recursive = TRUE, showWarnings = FALSE)
writeLines(c("# Copyright (c) 2020", "# Author: Jane Doe", "", "a <- 1"),
           file.path(project, "R", "a.R"))
writeLines(c("# Copyright (c) 2019", "# Author: Ada Lovelace", "", "b <- 2"),
           file.path(project, "R", "b.R"))

stamp_update(project, edits, recursive = TRUE)
```

Every stamped file in the tree now carries the new year and author, from a
single call.

## Preview and back up

Updating is editing, so the same `action` choices from stamping apply.
`action = "dryrun"` shows what would change without touching the file:

```{r}
h <- tempfile(fileext = ".R")
writeLines(c("# Copyright (c) 2019", "# Author: Ada Lovelace", "", "y <- 2"), h)
stamp_bump_year(h, action = "dryrun")
```

`action = "backup"` writes a `.bck` copy before editing.

## When a field is not there

The update functions only touch fields they can find. If you ask to update a
field the header does not contain, filestamp warns instead of silently doing
nothing, so a typo or a missing field does not pass unnoticed:

```{r}
k <- tempfile(fileext = ".R")
writeLines(c("# Copyright (c) 2021", "# Author: Grace Hopper", "", "z <- 3"), k)
stamp_update(k, license = "MIT")
```

The file is left unchanged, because there was no license line to update. Add one
by re-stamping with a template that includes it, or edit the file directly.
